Report on Late Winter2013 Production: Image Differencing\vspace{-2ex}

### 2013-04-09 W13report.tex
Incorporate many of Paul's comments

\usepackage[paperwidth=8.5in,paperheight=11in,centering,margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsbsy}
\usepackage{natbib}
\usepackage{rotating}
\input epsf
\usepackage{amsmath}
\usepackage{caption}
\usepackage{wasysym}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
%\usepackage{ulem}
%\usepackage{epstopdf}
\usepackage[toc,page]{appendix}

@@

We investigated the complexity of the basis functions used to model
the PSF--matching kernel, and the degree to which this kernel varies
across the images.  We generate several goodness of fit statistics for
a control sample of sources that were {\it not} used in the kernel fit
notably the mean squared error, in order to assess the interpolation ability
of the matching kernel.  These metrics are shown to correlate with the
numbers of false positives.  We calculate a mean squared error metric
that shows a minimum value close to the configurations that yielded
the lowest number of false positives, and shows promise as a means to
algorithmically choose the optimal pipeline configuration.

We compared the measured numbers of false detections to the number we
expect due to random fluctuations in a background Gaussian field.  The
latter is a function of both seeing in the images, and of the
detection threshold used.  At the fiducial detection threshold of
5--sigma, the number of expected false positives per LSST sensor is
3--12, depending upon seeing.  We measure a similar number in the
difference images, indicating that we are able to achieve the
systematic floor in false detection rate.  We examined our measured
false detection rate as a function of detection threshold and find
that it scales as expected.  From these measures, we estimate that in
order for there to be less than 1 expected false detection per sensor
(i.e.\ $\sim$20\% of the alert production rate) we must operate at a
detection threshold of 5.5--sigma (for seeings as good as 0.6''). This
result is seeing dependent.

We found a significant dependence of the false positive rate on our
understanding of the variance; even a 2\% misestimate of the noise may
change the false positive rate by a factor of 1.5--2 at 5--sigma.
Prefiltering is preferred in this regard, as our variance planes are
shown to track the variance to higher fidelity at the detection stage.
We find that over/undersubtraction of the image background leads to a
bias in the positive to negative false detection ratio, but not on the
overall number of detections.  Prefiltering may obviate the need for
deconvolution of the template image when the science image has better
seeing, which is a clear advantage of the technique.

@@

The basis functions themselves $K_i(u,v)$ are a degree of freedom in
this problem.  The original implementations \citep{Alard98,Alard00}
used a set of 3 Gaussians (represented by our {\tt Config} variable
nGauss), each with a different sigma (sigGauss), and each modified by
a Laguerre polynomial to a given order (degGauss).  Subsequent studies
\citep[e.g.][]{2007AN....328...16I} have suggested that a constant
ratio be maintained between the different Gaussian widths, such that
$\sigma_{i+1} = \beta \times \sigma_{i}$.  We use the value $\beta =
2.0$ for this production.  We set the overall scale for sigma by
noting that, under the assumption that the PSFs of the images are
Gaussian ($\sigma_S$ for the science image and $\sigma_T$ for the
template image), the sigma of the matching kernel should be simply
$\sigma_K^2 = \sigma_S^2 - \sigma_T^2$.  We use this width for the
central Gaussian in the basis sequences where we use more than one
Gaussian, or the sole Gaussian in the sequences when we use only one.

@@

In this production, we investigated two orders of operations for the
optimal filtering by the PSF.  In the first method, called
prefiltering, we PSF--match the template image to a pre--filtered
science image:
\begin{eqnarray}
D_{Pre}(x,y) & = & S(x,y) \circ PSF_S(u,v;x,y) - T(x,y) \otimes K(u,v;x,y) \nonumber \\ 
D'_{Pre}(x,y) & = & D_{Pre}(x,y). \nonumber 
\end{eqnarray}
By pre--smoothing the image that we are matching the template to --
$S(x,y) \circ PSF_S(u,v;x,y)$ -- we more frequently avoid the case of
PSF--matching {\it deconvolution}, where the FWHM of $T$ is broader
than $S$.  This pre--filtering increases $\sigma_S$ by $\sqrt{2}$.
The PSF matching kernels will need to be correspondingly larger to
account for the larger effective PSFs.  We are able to run detection
directly on this image.

In the second, which is the classical postfiltering implementation of
the technique, we create a difference image that is then
post--filtered with the PSF for detection:
\begin{eqnarray}
D_{Post}(x,y) & = & S(x,y) - T(x,y) \otimes K(u,v;x,y) \nonumber \\ 
D'_{Post}(x,y) & = & D_{Post}(x,y) \circ PSF_S(u,v;x,y).  \nonumber 
\end{eqnarray}

@@

Specialized simulations were developed with constrained astrophysical
complexity for the initial test data. The purpose of these images was
to define a baseline for how well image subtraction would perform in
the absence of chromatic effects (e.g.\ differential chromatic
refractions), and perturbations in the optical system. To accomplish
this the data comprised $i$-band simulations of a random distribution
of stars with a density of $\sim 1000$ stars per CCD and with an
equivalent $r$-band magnitude range ($19<r<21$). All stars in this
sample had identical spectral energy distributions
(km50\_5000.fits\_g20\_5140.gz), and, for all sources, variability,
proper motion, and parallax were set to zero. All simulations were
made at the zenith.

Nine sensors of the central raft were simulated, for a total of 27
difference images per configuration to use in our comparative
analysis.  Each simulation run was generated using the following
configurations:
\begin{itemize}
\item Visit 6866600: A pair of 300s exposure template images with
  an atmospheric seeing value of 0.88 arcsec.
\item Visit 6866601: A pair of 15s exposure exposures with 0.6
 arcsec atmospheric seeing.
\item Visit 6866602: A pair of 15s exposure exposures with 0.88
 arcsec atmospheric seeing.
\item Visit 6866603: A pair of 15s exposure exposures with 1.2
 arcsec atmospheric seeing.
\end{itemize}

For each configuration, v3.2.5 of the LSST phosim package was used to
generate the data. The sky background was dark and set to produce
$\sim$150 counts per pixel, with airglow variations set to
zero. Telescope and camera perturbations, tracking errors, field
rotation, and saturation and blooming effects were all disabled. To
enable the use of eimages directly within the analysis (i.e.\ to
remove the need for the construction of a full CCD from the amplifier
images, calibration of the simulated data, and the generation of
photometric zero points) the pixel quantum efficiency variations were
set to zero, cloud screens were turned off and no cosmic rays were
generated.  The phosim commands file for the generation of the
simulated data is listed in Appendix~\ref{sec-phosim}.  Reference
catalogs and astrometry.net indices were constructed from the input
catalogs.  A secondary set of simulation images was generated for an
extended Summer 2013 analysis.  These data are described in
Appendix~\ref{sec-phosimextend}.

@@

Major tasks that were completed in the run--up to production include:
implementing prefiltering for image subtraction (ticket \#2551);
implementing source selection for image subtraction (\#2554, \#2596);
implementing measurement for difference image source ({\tt DiaSource})
characterization (\#2595); implementing RegisterTask for
image--to--image registration (\#2562, \#2650); defining and
implementing metrics for testing ip\_diffim (\#2636); designing a tool
to scrape metadata from data repositories (\#2640); and implementing
flux measurement for PSF--filtered sources (\#2754).

@@

Next, we used the DiaCatalogSourceSelector to query the ImSim
reference catalog for appropriate sources to use for PSF matching.
This selector allows the user to specify the brightness and color
range of the objects, toggle star or galaxy selection, and to include
variable objects or not.  We selected a fraction (1/5) of these
objects to serve as a control sample, in order to assess the
effectiveness of $K(u,v;x,y)$ to subtract objects that were {\it not}
used in the least squares fit.

@@

%\subsubsection{Pipeline optimization}

 The first step is to use the
PSF Gaussian sigmas to define the {\it sizes} of the Gaussians shapes
in $K_i(u,v)$.  As mentioned above, the main width was chosen to be
$\sqrt{\sigma_S^2 - \sigma_T^2}$.  We investigated variations on the
configuration where we used one Gaussian with the above width, and
where we used three Gaussians, the central of which has this width and
the others being a factor of $\beta = 2.0$ smaller and larger.  We
modified each of the Gaussians by a set of Laguerre polynomials of a
given order.  The smallest Gaussian was modified to the specified
order, with the others modified by order/2.  Figure~\ref{basis}
provides an example basis set with this configuration.  The total
number of shapes in each basis were: $\sum_i^{\rm nGauss} ({\rm
  degGauss}_i+1)\times({\rm degGauss}_i+2)/2$ \footnote{ A notable
  exception to this algorithm is found in the case where $\sigma_T >
  \sigma_S$, i.e. a sharpening must be done to the template to match
  it to the science image.  This is the case of deconvolution.  A
  theoretical prescription for basis design in this case is given in
  \cite{0266-5611-26-8-085002}.  In our implementation of this algorithm, nGauss is {\it
    always} fixed to value 3, and degGauss is always fixed to 3.  The
  widths of the Gaussians are determined through the algorithm
  specified in \cite{0266-5611-26-8-085002}, using as inputs the
  sequence of Gaussians that we would have used to match a Gaussian of
  width $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convolved
  the science image and not the template image).  There are a variety
  of reasons that actually operating on the science image is
  undesirable, including: the science image is lower S/N than the
  template; the science image is likely to have masked defects,
  including cosmic rays and bad pixels, whose reach would be expanded
  due to the convolution; unmasked features such as faint cosmic rays
  are more likely to look PSF--like after convolution (cosmic rays
  will likely have been sigma clipped out of the coadd).  However, it
  is not clear that deconvolving the template image is in principle
  any better.  A far better option, which we explore here, is to
  pre--filter the science image, which will in the vast majority of
  cases provide a larger PSF (by a factor of $\sqrt{2}$) to match the
  template to.  This is the case for v6866601 in this study, where the
  postfiltered analysis yields deconvolution of the template, while
  the prefiltered analysis does not.  }.
The dimensions of the PSF matching kernel were chosen to be 6 times
the largest Gaussian width, with a minimum kernelSize of 21x21 pixels.
For kernels that are significantly smaller than this, the Gaussians
have significant (non--zero) power at the kernel boundaries, leading
to square systematic artifacts at the scale of the kernel in the
difference images (see Figure~\ref{fig_galleryBad}).

\begin{figure}
\centering{\includegraphics[width=0.8\textwidth]{figures/basis3.eps}}
\caption{Example basis set with nGauss=3 (highlighted in red),
  sigGauss=(0.5 sig, 1.0 sig, 2.0 sig), and degGauss=4, yielding
  Laguerre polynomials of order (4,2,2).}
\label{basis}
\end{figure}

For each source $j$ returned by DiaCatalogSourceSelector, we created a
{\tt KernelCandidate} instance, which holds substamps of $T_j(x,y)$
and $S_j(x,y)$ centered on the source.  These were used to fit for a
local solution $K_j(u,v)$; the ensemble of local solutions at each
source's $x,y$ was used to fit for the full spatial solution
$K(u,v;x,y)$.  We emphasize that the dimensions of the substamps are
important.  Note from equation~\ref{eq-soln} that the coefficients
$a_i$ are derived from a convolution of the template substamp with the
kernel basis functions $K_i$.  This convolution means that
kernelSize//2 pixels along the border of each substamp are rendered unusable.
To maintain a significant number of pixels in $C_i$, we set the {\tt
  KernelCandidate} dimensions to be twice that of the kernel (i.e. the
number of pixels remaining in $C_i$ is equal to the number of pixels
in the kernel, which is fewer than the number of bases $i$).  For each
{\tt KernelCandidate}, we solved for $K_j(u,v)$ and create a local
difference image $D_j(x,y)$.  We evaluated several statistics on this
local difference image, normalized by the square root of its variance,
which puts the pixels in units of sigma (see also
Section~\ref{sec-stats}).  We measured the mean sigma, the RMS of this
distribution, the $\chi^2$ of these pixels normalized by the degrees
of freedom (number of pixels in $D_j(x,y)$ minus the number of kernel
bases), and the mean squared error.  These are referred to as the
LOCAL kernel metrics, and reflect how appropriate the chosen kernel
basis is (see again Figure~\ref{fig_galleryBad}).

The ensemble of {\tt KernelCandidates} was then used to constrain a
spatial model of the kernel $K(u,v;x,y)$.  This was done using the
{\tt SpatialCellSet} formalism, whereby the {\tt KernelCandidates} are
distributed across the image in a grid of {\tt SpatialCells}.  This
formalism attempts to distribute constraints on the spatial model
evenly across the image.  At most 3 {\tt KernelCandidates} per {\tt
  SpatialCell} were input to the spatial kernel fit.  For the spatial
kernel model, we assumed that each of the kernel coefficients $a_i$
may be represented by a $N^{th}$ order 2--dimensional Chebyshev.
Normally, we would perform sigma--clipping iterations on candidates
having bad residuals after spatial modeling.  However in this
production we expect that all objects should fit, so we performed no
sigma clipping.  Having generated the full spatial solution
$K(u,v;x,y) = \sum_i a_i(x,y) K_i(u,v)$, we evaluated the kernel
solution at the position of each {\tt KernelCandidate}, created a
difference image using this kernel, and recalculated the metrics
defined above.  These are referred to as the SPATIAL kernel metrics.
Importantly, we also interpolated this solution to the positions of
the control sample, which were explicitly {\it not} used in the
spatial fit, and evaluate these same metrics.  The differences between
the {\tt KernelCandidate} LOCAL and SPATIAL metrics, as well as the
values of the SPATIAL metrics for the control sample, reflect the
appropriateness of the spatial model and our ability to interpolate
and extrapolate to the full extent of the images.  We used the Mean
Square Error (MSE) of the control sample to example the tradeoff
between bias and variance in the spatial interpolation.

We next ran the SourceDetectionTask on the difference image, using a
``polarity'' of {\it both} to search for both positive and negative
deviations.  We initially used a nominal detection threshold of 5
sigma.  In the case of pre--filtering, we detected on the difference
image directly.  In post--filtering, we convolved the difference
image with its PSF (by design, the same as the PSF of $S$) before
detection.  Since the fundamental image operations are the same, the images immediately preceding the
detection step ($D'$) should be {\it exactly} the same.  Detections
were merged using a grow radius of 2 pixels.

Finally, we ran DipoleMeasurementTask on the detections to create {\tt
  DiaSources}.  DipoleMeasurementTask is an extension of
SourceMeasurementTask that includes specific dipole measurement
algorithms.  Measurement was necessarily different in the
pre--filtered and post--filtered data.  In particular, in
post--filtering, detection happens on the filtered maximum likelihood
image and measurement on the associated pixels in the unfiltered
difference image.  In pre--filtering, the latter image does not exist,
and thus detection and measurement happen on the same image.  This
restricts the types of measurements that may be made on the sources
(see Section~\ref{sec-meas}).  While these measurements were
straightforward to implement, additional work needs to be done on how
to implement the remainder of the measurement suite on post--filtered
data (e.g. shapes, moments, and PSF goodness--of--fit metrics).

@@

We measured the flux of {\tt DiaSources} on postfiltered difference images
using PSF--weighted flux photometry (as implemented in the LSST
measurement algorithm "flux.psf").  For measuring fluxes of point
sources on prefiltered images we implemented a new algorithm
"flux.filtered".

We start with the definition of PSF--weighted flux:
%
\[flux = \frac{\sum_{(x,y)}I(x,y) * \phi(x,y)}{\sum_{(x',y')}\phi^2(x',y')}\]
%
where $I(x,y)$ is the image before PSF filtering and $\phi(x,y)$ is an
image of the PSF centered on the source. This equation assumes an
isolated point source, an assumption that is reasonable for difference
images.  We do not weight by the per--pixel variance to avoid
non--linearity in the flux measurement, at the expense of less precise
measurements.

Filtering is performed by convolving the image with its PSF, so the
peak of the filtered image of a given source is:
%
\[Ipeak_{filt} = \sum_{(x,y)}I(x,y) * \phi(x,y).\]
%
Hence the equation for measuring its flux reduces to:
%
\[flux = \frac{Ipeak_{filt}}{\sum_{(x',y')}\phi^2(x',y')}.\]

@@

The first statistic, KCDiffimStDev, measures the standard deviation of
all the non--{\tt EDGE} pixels in the difference image, normalized by
the sqrt of the variance of the difference image.  In the idealized
case, this value should be 1.0, i.e. pixels in the image are drawn
from a normal (0,1) distribution.  In cases of overfitting or
underfitting, this value will be lower than / higher than 1.0.

We also calculate a reduced $\chi^2_\nu$ value called KCDiffimChiSq.
The numerator of this term is the sum of the squares of the
noise--normalized pixel values.  This sum is normalized by the number
of pixels in the data, minus the number of basis shapes in the kernel
plus 1.

>>>>

\usepackage[paperwidth=8.5in,paperheight=11in,centering,margin=1in]{geometry}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amsbsy}
\usepackage{natbib}
\usepackage{rotating}
\input epsf
\usepackage{amsmath}
\usepackage{caption}
\usepackage{wasysym}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
%\usepackage{ulem}
%\usepackage{epstopdf}
\usepackage[toc,page]{appendix}

@@

We investigated the complexity of the basis functions used to model
the PSF--matching kernel, and the degree to which this kernel varies
across the images.  We generate several goodness of fit statistics for
a control sample of sources that were {\it not} used in the kernel fit
(notably the mean squared error), in order to assess the interpolation ability
of the matching kernel.  These metrics are shown to correlate with the
numbers of false positives.  We calculate a mean squared error metric
that shows a minimum value close to the configurations that yielded
the lowest number of false positives, and shows promise as a means to
algorithmically choose the optimal pipeline configuration.

We compared the measured numbers of false detections to the number we
expect due to random fluctuations in a background Gaussian field.  The
latter is a function of both seeing in the images, and of the
detection threshold used.  At the fiducial detection threshold of
5--sigma, the number of expected false positives per LSST sensor is
3--12, depending upon seeing (see Section \ref{sec-analyticfp}).  
We measure a similar number in the
difference images, indicating that we are able to achieve the
systematic floor in false detection rate.  We examined our measured
false detection rate as a function of detection threshold and find
that it scales as expected.  From these measures, we estimate that in
order for there to be less than 1 expected false detection per sensor
(i.e.\ $\sim$20\% of the alert production rate) we must operate at a
detection threshold of 5.5--sigma (for seeings as good as 0.6''). This
result is seeing dependent.

We found a significant dependence of the false positive rate on our
understanding of the variance; even a 2\% misestimate of the noise may
change the false positive rate by a factor of 1.5--2 at 5--sigma.
Pre--filtering is preferred in this regard, as our variance planes are
shown to track the variance to higher fidelity at the detection stage.
We find that over/under subtraction of the image background leads to a
bias in the positive to negative false detection ratio, but not on the
overall number of detections.  Pre--filtering may obviate the need for
deconvolution of the template image when the science image has better
seeing, which is a clear advantage of the technique.

@@

The basis functions themselves $K_i(u,v)$ are a degree of freedom in
this problem.  The original implementations \citep{Alard98,Alard00}
used a set of 3 Gaussians (represented by our {\tt Config} variable
nGauss), each with a different width (sigGauss), and each modified by
a Laguerre polynomial to a given order (degGauss).  Subsequent studies
\citep[e.g.][]{2007AN....328...16I} have suggested that a constant
ratio be maintained between the different Gaussian widths, such that
$\sigma_{i+1} = \beta \times \sigma_{i}$.  We use the value $\beta =
2.0$ for this production.  We set the overall scale for sigma by
noting that, under the assumption that the PSFs of the images are
Gaussian ($\sigma_S$ for the science image and $\sigma_T$ for the
template image), the sigma of the matching kernel should be simply
$\sigma_K^2 = \sigma_S^2 - \sigma_T^2$.  We use this width for the
central Gaussian in the basis sequences where we use more than one
Gaussian, or the sole Gaussian in the sequences when we use only one.

@@

In this production, we investigated two orders of operations for the
optimal filtering by the PSF.  In the first method, called
pre--filtering, we PSF--match the template image to a pre--filtered
science image:
\begin{eqnarray}
D_{Pre}(x,y) & = & S(x,y) \circ PSF_S(u,v;x,y) - T(x,y) \otimes K(u,v;x,y) \nonumber \\ 
D'_{Pre}(x,y) & = & D_{Pre}(x,y). \nonumber 
\end{eqnarray}
By pre--smoothing the image that we are matching the template to --
$S(x,y) \circ PSF_S(u,v;x,y)$ -- we more frequently avoid the case of
PSF--matching {\it deconvolution}, where the full width at half maximum (FWHM)
 of $T$ is broader
than $S$.  This pre--filtering increases $\sigma_S$ by a factor of $\sqrt{2}$.
The PSF matching kernels will need to be correspondingly larger to
account for the larger effective PSFs.  We are able to run detection
directly on this image with different measurement algorithms (see Section \ref{sec-meas}).

In the second, which is the classical post--filtering implementation of
the technique, we create a difference image that is then
post--filtered with the PSF for detection:
\begin{eqnarray}
D_{Post}(x,y) & = & S(x,y) - T(x,y) \otimes K(u,v;x,y) \nonumber \\ 
D'_{Post}(x,y) & = & D_{Post}(x,y) \circ PSF_S(u,v;x,y).  \nonumber 
\end{eqnarray}

@@

Specialized simulations were developed with constrained astrophysical
complexity for the initial test data. The purpose of these images was
to define a baseline for how well image subtraction would perform in
the absence of chromatic effects (e.g.\ differential chromatic
refractions), and perturbations in the optical system. To accomplish
this the data comprised $i$-band simulations of a random distribution
of stars with a density of $\sim 1000$ stars per CCD and with an
equivalent $r$-band magnitude range ($19<r<21$). All stars in this
sample had identical spectral energy distributions
(km50\_5000.fits\_g20\_5140.gz; see the phosim page,{\tt http://ls.st/1vo} for links to the
spectra), and, for all sources, variability,
proper motion, and parallax were set to zero. All simulations were
made at the zenith.

Nine sensors of the central raft were simulated, for a total of 27
difference images per configuration to use in our comparative
analysis.  In each case a pair of snap exposures was generated, however only
the first snap was used in the analysis.  Each simulation run was generated using the following
configurations:
\begin{itemize}
\item Visit 6866600: A pair of 300s exposure template images with
  an atmospheric seeing value of 0.88 arcsec.
\item Visit 6866601: A pair of 15s exposure exposures with 0.6
 arcsec atmospheric seeing.
\item Visit 6866602: A pair of 15s exposure exposures with 0.88
 arcsec atmospheric seeing.
\item Visit 6866603: A pair of 15s exposure exposures with 1.2
 arcsec atmospheric seeing.
\end{itemize}

For each configuration, v3.2.5 of the LSST phosim package (see {\tt http://ls.st/1vo} for usage)
 was used to
generate the data. The sky background was dark and set to produce
$\sim$150 counts per pixel, with airglow variations set to
zero. Telescope and camera perturbations, tracking errors, field
rotation, and saturation and blooming effects were all disabled. To
enable the use of eimages directly within the analysis (i.e.\ to
remove the need for the construction of a full CCD from the amplifier
images, calibration of the simulated data, and the generation of
photometric zero points) the pixel quantum efficiency variations were
set to zero, cloud screens were turned off and no cosmic rays were
generated.  The phosim commands file for the generation of the
simulated data is listed in Appendix~\ref{sec-phosim}.  Reference
catalogs and astrometry.net indices were constructed from the input
catalogs.  A secondary set of simulation images was generated for an
extended Summer 2013 analysis.  These data are described in
Appendix~\ref{sec-phosimextend}.

@@

Major tasks that were completed in the run--up to production include:
implementing pre--filtering for image subtraction (ticket \#2551);
implementing source selection for image subtraction (\#2554, \#2596);
implementing measurement for difference image source ({\tt DiaSource})
characterization (\#2595); implementing RegisterTask for
image--to--image registration (\#2562, \#2650); defining and
implementing metrics for testing ip\_diffim (\#2636); designing a tool
to scrape metadata from data repositories (\#2640); and implementing
flux measurement for PSF--filtered sources (\#2754).

@@

Next, we used the DiaCatalogSourceSelector to query the ImSim
reference catalog for appropriate sources to use for PSF matching.
This selector allows the user to specify the brightness and color
range of the objects, toggle star or galaxy selection, and to include
variable objects or not.  In this semi-ideal case we use all sources.
We selected a fraction (1/5) of these
objects to serve as a control sample, in order to assess the
effectiveness of $K(u,v;x,y)$ to subtract objects that were {\it not}
used in the least squares fit.

@@

%\subsubsection{Pipeline optimization}
\begin{figure}[!hpb]
\centering{\includegraphics[width=0.8\textwidth]{figures/basis3.eps}}
\caption{Example basis set with nGauss=3 (highlighted in red),
  sigGauss=(0.5 sig, 1.0 sig, 2.0 sig), and degGauss=4, yielding
  Laguerre polynomials of order (4,2,2).}
\label{basis}
\end{figure}

 The first step is to use the
PSF Gaussian widths to define the {\it sizes} of the Gaussians' shapes
in $K_i(u,v)$.  As mentioned above, the main width was chosen to be
$\sqrt{\sigma_S^2 - \sigma_T^2}$.  We investigated variations on the
configuration where we used one Gaussian with the above width, and
where we used three Gaussians, the central of which has this width and
the others being a factor of $\beta = 2.0$ smaller and larger.  We
modified each of the Gaussians by a set of Laguerre polynomials of a
given order.  The smallest Gaussian was modified to the specified
order, with the others modified by floor(order/2).  Figure~\ref{basis}
provides an example basis set with this configuration.  The total
number of shapes in each basis were: $\sum_i^{\rm nGauss} ({\rm
  degGauss}_i+1)\times({\rm degGauss}_i+2)/2$ \footnote{ A notable
  exception to this algorithm is found in the case where $\sigma_T >
  \sigma_S$, i.e. a sharpening must be done to the template to match
  it to the science image.  This is the case of deconvolution.  A
  theoretical prescription for basis design in this case is given in
  \cite{0266-5611-26-8-085002}.  In our implementation of this algorithm, nGauss is {\it
    always} fixed to value 3, and degGauss is always fixed to 3.  The
  widths of the Gaussians are determined through the algorithm
  specified in \cite{0266-5611-26-8-085002}, using as inputs the
  sequence of Gaussians that we would have used to match a Gaussian of
  width $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convolved
  the science image and not the template image).  There are a variety
  of reasons that actually operating on the science image is
  undesirable, including: the science image is lower S/N than the
  template; the science image is likely to have masked defects,
  including cosmic rays and bad pixels, whose reach would be expanded
  due to the convolution; unmasked features such as faint cosmic rays
  are more likely to look PSF--like after convolution (cosmic rays
  will likely have been sigma clipped out of the coadd).  However, it
  is not clear that deconvolving the template image is in principle
  any better.  A far better option, which we explore here, is to
  pre--filter the science image, which will in the vast majority of
  cases provide a larger PSF (by a factor of $\sqrt{2}$) to match the
  template to.  This is the case for v6866601 in this study, where the
  post--filtered analysis yields deconvolution of the template, while
  the pre--filtered analysis does not.  }.
The dimensions of the PSF matching kernel were chosen to be 6 times
the largest Gaussian width, with a minimum kernelSize of 21x21 pixels.
For kernels that are significantly smaller than this, the Gaussians
have significant (non--zero) power at the kernel boundaries, leading
to square systematic artifacts at the scale of the kernel in the
difference images (see Figure~\ref{fig_galleryBad}).
\begin{figure}[!hpb]
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/deconv2.eps} 
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/size2.eps} \\
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/order2.eps} 
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/shape2.eps} \\
\caption{Four example panels showing failure modes of the image
  subtraction software.  Each panel itself shows a 2x2 pair of
  {\tt KernelCandidate} difference images; on the {\it left} subpanel is the residual from
  the LOCAL fitting, and on the {\it right} from the SPATIAL fitting.
  Starting in the upper left and going clockwise, the first panel
  shows the residuals when deconvolving the template image; the
  ringing is characteristic of the deconvolution process.  Next, a set
  of difference images where the kernel dimensions (kernelSize) are
  too small to match the sources, leaving square kernel--sized
  residuals around each object.  Next, a set of images where the
  kernel shape (sigGauss) is inappropriate to match the sources; note
  that the LOCAL residuals show circular residuals, indicating the
  basis set itself is at fault (as opposed to the spatial model).
  Finally, in the lower--left, a set of images that demonstrate how
  the residuals degrade when the spatial kernel model is at fault.
  Note that the LOCAL residuals appear to be white noise, while the
  SPATIAL residuals show a clear dipole signature.  For this reason, a
  comparison between the LOCAL and SPATIAL residuals is a useful
  diagnostic of the spatial model.  }
\label{fig_galleryBad}
\end{figure}

For each source $j$ returned by DiaCatalogSourceSelector, we created a
{\tt KernelCandidate} instance, which holds substamps of $T_j(x,y)$
and $S_j(x,y)$ centered on the source.  These were used to fit for a
local solution $K_j(u,v)$; the ensemble of local solutions at each
source's $x,y$ was used to fit for the full spatial solution
$K(u,v;x,y)$.  We emphasize that the dimensions of the substamps are
important.  Note from equation~\ref{eq-soln} that the coefficients
$a_i$ are derived from a convolution of the template substamp with the
kernel basis functions $K_i$.  This convolution means that
kernelSize//2 pixels along the border of each substamp are rendered unusable.
To maintain a significant number of pixels in $C_i$, we set the {\tt
  KernelCandidate} dimensions to be twice that of the kernel (i.e. the
number of pixels remaining in $C_i$ is equal to the number of pixels
in the kernel, which is fewer than the number of bases $i$).  For each
{\tt KernelCandidate}, we solved for $K_j(u,v)$ and create a local
difference image $D_j(x,y)$.  We evaluated several statistics on this
local difference image, normalized by the square root of its variance,
which puts the pixels in units of $\sigma$ (see also
Section~\ref{sec-stats}).  We measured the mean sigma, the RMS of this
distribution, the $\chi^2$ of these pixels normalized by the degrees
of freedom (number of pixels in $D_j(x,y)$ minus the number of kernel
bases), and the mean squared error.  These are referred to as the
LOCAL kernel metrics, and reflect how appropriate the chosen kernel
basis is (see again Figure~\ref{fig_galleryBad}).

The ensemble of {\tt KernelCandidates} was then used to constrain a
spatial model of the kernel $K(u,v;x,y)$.  This was done using the
{\tt SpatialCellSet} formalism, whereby the {\tt KernelCandidates} are
distributed across the image in a grid of {\tt SpatialCells}.  This
formalism attempts to distribute constraints on the spatial model
evenly across the image.  At most 3 {\tt KernelCandidates} per {\tt
  SpatialCell} were input to the spatial kernel fit.  For the spatial
kernel model, we assumed that each of the kernel coefficients $a_i$
may be represented by a $N^{th}$ order 2--dimensional Chebyshev.
Normally, we would perform sigma--clipping iterations on candidates
having bad residuals after spatial modeling.  However in this
production we expect that all objects should fit, so we performed no
sigma clipping.  Having generated the full spatial solution
$K(u,v;x,y) = \sum_i a_i(x,y) K_i(u,v)$, we evaluated the kernel
solution at the position of each {\tt KernelCandidate}, created a
difference image using this kernel, and recalculated the metrics
defined above.  These are referred to as the SPATIAL kernel metrics.
Importantly, we also interpolated this solution to the positions of
the control sample, which were explicitly {\it not} used in the
spatial fit, and evaluate these same metrics.  The differences between
the {\tt KernelCandidate} LOCAL and SPATIAL metrics, as well as the
values of the SPATIAL metrics for the control sample, reflect the
appropriateness of the spatial model and our ability to interpolate
and extrapolate to the full extent of the images.  We used the Mean
Square Error (MSE) of the control sample to examine the tradeoff
between bias and variance in the spatial interpolation.

We next ran the SourceDetectionTask on the difference image, using a
``polarity'' of {\it both} to search for both positive and negative
deviations.  We initially used a nominal detection threshold of 5
$\sigma$.  In the case of pre--filtering, we detected on the difference
image directly.  In post--filtering, we convolved the difference
image with its PSF (by design, the same as the PSF of $S$) before
detection.  Since the fundamental image operations are the same, the images immediately preceding the
detection step ($D'$) should be {\it exactly} the same.  Detections
were merged using a grow radius of 2 pixels.

Finally, we ran DipoleMeasurementTask on the detections to create {\tt
  DiaSources}.  DipoleMeasurementTask is an extension of
SourceMeasurementTask that includes specific dipole measurement
algorithms.  Measurement was necessarily different in the
pre--filtered and post--filtered data.  In particular, in
post--filtering, detection happens on the filtered maximum likelihood
image and measurement on the associated pixels in the unfiltered
difference image.  In pre--filtering, the latter image does not exist,
and thus detection and measurement happen on the same image.  This
influences how measurements made on the sources are interpreted
(see Section~\ref{sec-meas}).  While these measurements were
straightforward to implement, additional work needs to be done on how
to implement the remainder of the measurement suite on post--filtered
data (e.g. shapes, moments, and PSF goodness--of--fit metrics).

@@

We measured the flux of {\tt DiaSources} on post--filtered difference images
using PSF--weighted flux photometry (as implemented in the LSST
measurement algorithm ``flux.psf").  For measuring fluxes of point
sources on pre--filtered images we implemented a new algorithm
``flux.filtered".

We start with the definition of PSF--weighted flux assuming a normalized psf, $\phi$:
%
\[flux = \frac{\sum_{(x,y)}I(x,y) * \phi(x,y)}{\sum_{(u,v)}\phi^2(u,v)}\]
%
where $I(x,y)$ is the image before PSF filtering and $\phi(x,y)$ is an
image of the PSF centered on the source. This equation assumes an
isolated point source, an assumption that is reasonable for difference
images.  We do not weight by the per--pixel variance to avoid
non--linearity in the flux measurement, at the expense of less precise
measurements at the bright end.

Filtering is performed by convolving the image with its PSF, so the
peak of the filtered image of a given source is:
%
\[max(I_{filt}) = \sum_{(x,y)}I(x,y) * \phi(x,y).\]
%
Hence the equation for measuring its flux reduces to:
%
\[flux = \frac{max(I_{filt})}{\sum_{(x',y')}\phi^2(x',y')}.\]

@@

The first statistic, KCDiffimStDev, measures the standard deviation of
all the non--{\tt EDGE} pixels in the difference image, normalized by
the square root of the variance of the difference image.  In the idealized
case, this value should be 1.0, i.e. pixels in the image are drawn
from a normal (0,1) distribution.  In cases of overfitting or
underfitting, this value will be lower than / higher than 1.0.

We also calculate a reduced $\chi^2_\nu$ value called KCDiffimChiSq.
The numerator of this term is the sum of the squares of the
noise--normalized pixel values.  This sum is normalized by the number
of pixels in the data minus (the number of basis shapes in the kernel
plus 1).

Finally in order to investigate the bias--variance trade--off in the
overall fit, we calculate the mean square error (MSE) of each {\tt
  KernelCandidate's} difference image as statistic KCDiffimMseResids.
We define the bias as $\left| data - model \right|$, the variance as
$(data - model)^2$, and the MSE as {\tt bias$^2$ +
  variance}.  In this context, the bias is the mean of the difference
image, and the variance is the mean square of the difference image.
Ideally, the MSE gives a means for identifying the location in
parameter space of optimal fitting without overfitting.
